---
title: "Analytical Report"
author: "Kyle Fairbairn"
date: "2023/08/04"
---


## Introduction
This report presents the results of the analytical queries from the MySQL data warehouse for weekly analysis.

## Analytical Query I: Top Five Sales Reps with the Most Sales Broken Down by Year

### Query Description
The following query retrieves the top five sales reps with the most sales, broken down by year.

```{r InstallPackages}
# Packages used for Practicum II 
packages <- c("RMySQL","RSQLite", "DBI", "lubridate", "ggplot2")

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Packages loading
invisible(lapply(packages, library, character.only = TRUE))

```

```{r connectToRemoteDb}
connectToRemoteDb <- function () {
  
  db_name_fh <- "sql9637031"
  db_user_fh <- "sql9637031"
  db_host_fh <- "sql9.freemysqlhosting.net"
  db_pwd_fh <- "PASSWORD"
  db_port_fh <- 3306
  
  # Connect to remote server database
  mydb.fh <-  dbConnect(RMySQL::MySQL(), user = db_user_fh, password = db_pwd_fh,
                        dbname = db_name_fh, host = db_host_fh, port = db_port_fh)
  
  mydb <- mydb.fh
  
  return(mydb)
}

connectToLocalDb <- function() {
  mydb <- dbConnect(RSQLite::SQLite(), dbname = "sales.db")
  return (mydb)
}

mysqlDb <- connectToRemoteDb()
sqlLiteDb <- connectToLocalDb()
```

## Top Five Sales Reps 

## To summarize the report
### There is the sales reps name 
### There assigned location of sales to help determine the market


```{sql connection=mysqlDb}

SELECT 
    year,
    sales_rep_name,
    total_sold AS total_sales,
    region as Region
FROM 
    rep_facts
GROUP BY 
    year, sales_rep_name
ORDER BY 
    year, total_sales DESC
LIMIT 5;


```
## Analytical Query II: Total sold per year.

## Total Sales for the year 2020 by month 
### If a rep sold zero they wont be on graph 

```{r}

query <- "SELECT 
            year, 
            month, 
            sales_rep_name as Sales_Rep,
            total_monthly_sales AS total_monthly_sales
          FROM 
            rep_facts
          GROUP BY 
            year, month, sales_rep_name
          ORDER BY 
            year, month;"


result <- dbGetQuery(mysqlDb, query)

# Convert year and month columns to a single date column
result$date <- as.Date(paste(result$year, result$month, "01", sep = "-"))

ggplot(result, aes(x = date, y = total_monthly_sales, color = Sales_Rep)) +
  geom_line() +
  labs(title = "Total Monthly Sales per Sales Rep",
       x = "Date",
       y = "Total Monthly Sales") +
  scale_y_continuous(labels = scales::comma) +  # Format y-axis labels with commas
  theme_minimal()

```




```{r}
dbDisconnect(mysqlDb)
dbDisconnect(sqlLiteDb)
```


